The \dealii Library, Version 9.5

### 2023-05-23 9.5/paper.tex
Write the language standard section.

Mention the C++20 work. Starting point: \dealii{} relies heavily on
templates that are explicitly instantiated.

>>>>

\dealii{} currently uses C++14 as the language standard to which its
code base is written. It can also use the classes
\texttt{std::optional} and \texttt{std::variant} if the compiler
supports C++17, but falls back to implementations obtained via the
BOOST library otherwise, and this is true also for a number of
individual functions that were introduced in C++17 or C++20.

As part of the current release, \dealii{} now also uses some C++20
features to annotate classes and functions with regard to properties
template arguments need to satisfy. This aids in situations such as
with the
following function:
\begin{c++}
  template <class MeshType>
  std::vector<typename MeshType::active_cell_iterator>
  find_cells_adjacent_to_vertex(const MeshType &   mesh,
                                const unsigned int vertex);
\end{c++}
This function is intended to be called with either a
\texttt{Triangulation} or \texttt{DoFHandler} object as first
argument (and documents this requirement), but the compiler can not check this requirement at the call
site and will gladly call it with any other kind of object as
well. Because the implementation of the function is in a \texttt{.cc}
file, and the template is only instantiated for the two classes
mentioned above, calling the function (erroneously) with anything else
as first argument is not detected at compile time, but only later when
the linker reports an undefined symbol.

We have started to address this by annotating functions using
C++20-style ``\texttt{requires}'' clauses:
\begin{c++}
  template <class MeshType>
    requires (concepts::is_triangulation_or_dof_handler<MeshType>)
  std::vector<typename MeshType::active_cell_iterator>
  find_cells_adjacent_to_vertex(const MeshType &   mesh,
                                const unsigned int vertex)
\end{c++}
If the compiler supports C++20, then the added clause will cause the
compiler to reject any call to the function for which the first
argument \texttt{MeshType} does not satisfy the named concept, which
is internally defined as the type being either a \texttt{Triangulation} or \texttt{DoFHandler}
object, as intended. (The \texttt{requires} annotation is suppressed if the compiler
does not support C++20.)

Given the heavy dependence of \dealii{} on templates, there are likely
hundreds or thousands of locations that should be annotated with
requirements on template arguments over time; for the moment, the
library contains some 300 of these \texttt{requires} clauses.
